---
title: "ZZSC5855 --- Cluster analysis"
author:
- "Pavel Krivitsky"
- "Department of Statistics"
- "University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---


# Packages
```{r}
library(here)
library(GGally)
# stats is loaded by default
library(readr)
library(dplyr)
library(candisc)
library(cluster)
library(mclust)
```

# Iris Example
## Data
Suppose that we had mixed up our lab samples and no longer know which specimen comes from which species. Can we recover the species?
```{r}
data(iris)
iris0 <- iris[-5]
```

This plotting function will be helpful for visualising the clusterings produced:
```{r}
irisplot <- function(c){
# Plot the Iris data variables pairwise, with clusters identified by colour and species by symbol.
    pairs(iris0, col=c+1, pch=as.numeric(as.factor(iris$Species)))
# Alternative approach: find the canonical discriminants for separating newly detected clusters, and plot those:
    m <- lm(as.matrix(iris0)~factor(c)) # Fit a linear model with clustering as RHS.
    cd <- candisc(m)
    plot(cd, col=c("red","green","purple"))
}
```

## Hierarchical clustering

### Built-in (`stats` package)

We begin with hierarchical clustering performed by `hclust()`, using the Ward's method with squared Euclidean distance. The following code produces a complete hierarchy:
```{r}
iris.h <- hclust(dist(iris0), method="ward.D2")
plot(iris.h)
```
We then cut it at 3 clusters
```{r}
(iris.h3 <- cutree(iris.h,3)) # Three clusters
table(species=iris$Species, cluster=iris.h3) # Can it recover the species?
irisplot(iris.h3)
```
We observe 16 misclusterings, mostly virginicas being put into the majority-versicolor cluster.

The canonical discrimination plot identifies all the variables except for sepal width as being relatively redundant with each other for the purposes of this clustering.

### `cluster` package

The `cluster` package allows for more flexible clustering, as well as support for silhouette calculation:
```{r}
iris.h <- agnes(iris0, method="ward") # More flexible; see help.
irisplot(cutree(iris.h,3))
# Which cluster number gives the best average silhouette width?
d <- dist(iris0)
plot(2:10,sapply(2:10, function(k) summary(silhouette(cutree(iris.h,k),d))$avg.width))
plot(silhouette(cutree(iris.h,3), d))
```
The clustering is the same, but the silhouette plot does not select 3 clusters: in fact, it prefers 2, perhaps blurring together versicolor with virginica.

## Non-hierarchical clustering

### $k$-means

Now, let's try $k$-means from the `stats` package:
```{r}
iris.3means <- kmeans(iris0,3)
irisplot(iris.3means$cluster)
table(species=iris$Species, cluster=iris.3means$cluster)
plot(2:10,sapply(2:10, function(k) summary(silhouette(kmeans(iris0,k)$cluster,d))$avg.width))
plot(silhouette(iris.3means$cluster, d))
```
$k$-means has the same misclassification rate, again mostly virginicas being placed into majority-versicolor clusters.

### $k$-medioids

$k$-medioids (a.k.a. "Partitioning Around Medioids"):
```{r}
iris.3meds <- pam(iris0,3)
irisplot(iris.3meds$cluster)
table(species=iris$Species, cluster=iris.3meds$cluster)
plot(2:10,sapply(2:10, function(k) summary(silhouette(pam(iris0,k)$cluster,d))$avg.width))
plot(silhouette(iris.3meds$cluster, d))
```
Similar results from $k$-medioids.

### Model-based

Lastly, let's try to find the best model-based clustering:
```{r}
# Automatic best model and cluster count selection via BIC:
(iris.bestMBC <- Mclust(iris0))
plot(iris.bestMBC, what=c("BIC", "classification"))
```
The model selected is the VEV model: varying dispersions, common eccentricities, and varying orientations. Two clusters are selected, again blurring virginica with versicolor.

Alternatively, let's try forcing 3 clusters:
```{r}
(iris.3MBC <- Mclust(iris0,G=3))
plot(iris.3MBC, what=c("classification", "uncertainty", "density"))
irisplot(iris.3MBC$classification)
table(species=iris$Species, cluster=iris.3MBC$classification)
```
A VEV model is again selected. Forcing 3 clusters actually produces a significantly lower error rate than others, with only 5 versicolors clustered with the virginicas.


