---
title: "ZZSC5855 --- Cluster analysis"
author:
- Pavel Krivitsky
- Department of Statistics
- University of New South Wales
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---


# Packages
```{r}
library(here)
library(GGally)
# stats is loaded by default
library(readr)
library(dplyr)
library(candisc)
library(cluster)
library(mclust)
```


# Challenges

Recall the dataset `pizza.csv`, containing nutritional data from a variety of pizza brands:

`brand`
: Pizza brand (class label)

`id`
: Sample analysed

`prot`
: Amount of protein per 100 grams in the sample

`fat`
: Amount of fat per 100 grams in the sample

`ash`
: Amount of ash per 100 grams in the sample

`sodium`
: Amount of sodium per 100 grams in the sample

`carb`
: Amount of carbohydrates per 100 grams in the sample

`cal`
: Amount of calories per 100 grams in the sample

```{r}
pizza <- read_csv(here("datasets","pizza.csv"))
ggpairs(pizza, mapping=aes(col=brand, alpha=0.3))
```
Suppose that we have lost track of which pizza came from which brand, i.e., 
```{r}
pizza0 <- select(pizza, -brand)
```

### Task 1: Supposing that we know that there are 10 brands, try using the various clustering methods to group the pizzas by brand. How well can you recover them?


### Task 2: Now, suppose that we don't know how many brands there were. How many would you infer from the unlabelled data?

