---
title: "ZZSC5855 --- Multivariate LM and MANOVA Example"
author:
- "Pavel Krivitsky"
- "Department of Statistics"
- "University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---


# Packages
```{r}
library(here)
library(GGally)
library(dplyr)
library(readr)
library(purrr)
```

# Fitness Club Example
Three physiological and three exercise variables were measured on twenty middle aged men in a fitness club. Canonical correlation is used to determine if the physiological variables are related in any way to the exercise variables. 

## Data

```{r}
'weight waist pulse chins situps jumps
   191  36  50   5  162   60
   189  37  52   2  110   60
   193  38  58  12  101  101
   162  35  62  12  105   37
   189  35  46  13  155   58
   182  36  56   4  101   42
   211  38  56   8  101   38
   167  34  60   6  125   40
   176  31  74  15  200   40
   154  33  56  17  251  250
   169  34  50  17  120   38
   166  33  52  13  210  115
   154  34  64  14  215  105
   247  46  50   1   50   50
   193  36  46   6   70   31
   202  37  62  12  210  120
   176  37  54   4   60   25
   157  32  52  11  230   80
   156  33  54  15  225   73
   138  33  68   2  110   43' %>%
textConnection() %>% read.table(header=TRUE) -> fitness

fitness

```
## Exploratory data analysis

```{r}
summary(fitness)
ggpairs(fitness)
```
The data do have some outliers.

## Multivariate linear model analysis
### Estimation
```{r}
class(fitness.mlm <- lm(cbind(chins,situps,jumps)~waist+pulse+weight, data=fitness))
summary(fitness.mlm)
coef(fitness.mlm) # Now a p by k matrix.
vcov(fitness.mlm) # Now a p*k by p*k matrix.
estVar(fitness.mlm) # Now a p by p matrix.
resid(fitness.mlm) # Now a p-column matrix.
fitted(fitness.mlm) # Now a p-column matrix.
```

### Hypothesis testing
```{r}
# ? anova.mlm for options
anova(fitness.mlm) # Pillai's trace by default
anova(fitness.mlm, test="Wilks")
anova(fitness.mlm, test="Hotelling-Lawley")
anova(fitness.mlm, test="Roy")
# Sphericity test also possible---see help.

 # Test the effect of weight in the presence of the others:
fitness.mlm0 <- lm(cbind(chins,situps,jumps)~waist+pulse, data=fitness) 
anova(fitness.mlm, fitness.mlm0)
```
### Diagnostics
```{r, eval=FALSE}
plot(fitness.mlm) # Error: not implemented.
```
Pairwise plots of all residuals:
```{r}
ggpairs(as_tibble(resid(fitness.mlm))) # Note: type="pearson" doesn't work as of R 3.6!
```

Make our own Pearson residuals:
```{r}
# Residual standard deviations: square root of the diagonal of the estimated covariance of residuals:
sds <- fitness.mlm %>% estVar %>% diag %>% sqrt
PR <- sweep(resid(fitness.mlm), 2, sds, `/`)
ggpairs(as_tibble(PR))
```

Related: decorrelate the residuals:
```{r}
# Inverse of the Cholesky decomposition (square root) of the estimated variance-covariance matrix of the residuals
Uinv <- estVar(fitness.mlm) %>% chol %>% solve
(PRI <- resid(fitness.mlm)%*%Uinv) %>% cor %>% zapsmall # Now close to identity matrix.
ggpairs(as_tibble(PRI))
```

Residuals vs. fitted:
```{r}
# I.e., join the predictors with the residuals, and plot the predictors horizontally and residuals vertically:
pairs(as_tibble(cbind(PR,fitted(fitness.mlm))), horInd=1:3,verInd=4:6,
panel=function(x,y,...){abline(h=0,col="gray");points(x[abs(y)<2],y[abs(y)<2]);if(any(abs(y)>=2)) text(x[abs(y)>=2],y[abs(y)>=2],labels=which(abs(y)>=2));lines(lowess(x,y),col="red")})
```

