---
title: "ZZSC5855 --- Multivariate LM and MANOVA Example"
author:
- "Pavel Krivitsky"
- "Department of Statistics"
- "University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---


# Packages
```{r}
library(here)
library(GGally)
library(dplyr)
library(readr)
library(purrr)
```


# Challenge

Recall the dataset `pizza.csv`, containing nutritional data from a variety of pizza brands:

`brand`
: Pizza brand (class label)

`id`
: Sample analysed

`prot`
: Amount of protein per 100 grams in the sample

`fat`
: Amount of fat per 100 grams in the sample

`ash`
: Amount of ash per 100 grams in the sample

`sodium`
: Amount of sodium per 100 grams in the sample

`carb`
: Amount of carbohydrates per 100 grams in the sample

`cal`
: Amount of calories per 100 grams in the sample


We will focus on brand A:
```{r}
pizzaA <- read_csv(here("datasets","pizza.csv")) %>% filter(brand=="A") %>% select(-brand)
```

### Task 1: Fit and diagnose a linear model jointly predicting caloric, ash, and sodium content from fat, protein, and carbohydrate content.


### Task 2: Test whether carbohydrate content is predictive of the three response variables (jointly) in the presence of other predictors.


### Task 3: Test whether fat content is predictive of the three response variables (jointly) in the presence of other predictors.


### Task 4: Now, consider the full dataset, with brands. Use MANOVA to test whether different brands have different population mean measured pizza properties.

```{r}
pizza <- read_csv(here("datasets","pizza.csv"))
ggpairs(pizza, mapping=aes(col=brand, alpha=0.3))
```


