---
title: "ZZSC5855 --- Multivariate Normal Distribution"
author:
- "Pavel Krivitsky"
- "Deparment of Statistics"
- "University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---


# Packages

```{r, message=FALSE}
library(here)
library(mvtnorm)
library(GGally)
library(MVN)
library(dplyr)
library(readr)
```

# Multivariate normal random variables

## Generating

The `*mvnorm` family of functions from the `mvtnorm` can be used to calculate densities and generate values from the multivariate normal. We will use the `ggpairs` function found in package `GGally`:
```{r}
(mu <- c(2,3,4))
(Sigma <- diag(1:3) + 1)
x <- rmvnorm(1000, mu, Sigma)
dim(x)
```
we see that `x` has samples in rows and variables in columns.

Let's plot these variables:
```{r}
ggpairs(as.data.frame(x))
```

## Marginally normal vs. jointly normal

Consider two variables $Z_1=(2W-1)Y$ and $Z_2=Y$, where $Y\sim N(0,1)$ and, independently, $W \sim \operatorname{Binomial}(1,1/2)$ (so $2W-1$ takes $-1$ and $+1$ with equal probability).
```{r}
y <- rnorm(1000)
w <- rbinom(1000,1,1/2)
z <- cbind((2*w-1)*y, y)
ggpairs(as.data.frame(z))
```
We see that they are normal marginally yet clearly not jointly.

What happens when we use Mardia's diagnostics, using the `mvn` function in the `MVN` package?
```{r}
mvn(z)
mvn(z, mvnTest="mardia", univariateTest="SW") # Methods we've discussed.
```
We see that normality tests for individual dimensions (i.e., Shapiro--Wilk) pass, but Mardia's kurtosis does not.

Lastly, let us try viewing some *projections* of this distribution. For a true multivariate normal distribution, all projections are also normal. Here,
```{r}
zp1 <- z%*%c(1,-1)
plot(density(zp1))
```

Contrasting with a true MVN, even correlated,
```{r}
xp1 <- x%*%c(1,-1,0)
plot(density(xp1))
xp2 <- x%*%rnorm(3) # A random projection.
plot(density(xp2))
```

# Diagnosing multivariate normality

These data, taken from Johnson and Wichern (2007), consist of measurements of radiation emitted from 42 randomly chosen microwave ovens, with doors open and closed. Load the data and plot:
```{r}
ovens <- read_csv(here("datasets","ovens.csv"))
ggpairs(ovens)
```
Clearly, the data are not even close to MVN. Both variables are very strongly right-skewed.

Let's see the hypothesis tests:
```{r}
mvn(ovens)
mvn(ovens, mvnTest="mardia", univariateTest="SW") # Methods we've discussed.
```
We see that the multivariate tests for skewness and kurtosis reject the null hypotheses at any reasonable threshold.

Now, let's try transforming the data. It should be noted that transformations can change the interpretation of the results.
```{r}
ovens4 <- ovens^(1/4) # Take 4th root.
ggpairs(ovens4)
```
We now see far more normal-like shapes; though the `open` variable looks like it might be slightly bimodal or right-skewed.

Let's try the tests:
```{r}
mvn(ovens4)
mvn(ovens4, mvnTest="mardia", univariateTest="SW") # Methods we've discussed.
```
We see that both skewness and kurtosis are not inconsistent with a multivariate normal distribution.

