---
title: "ZZSC5855 --- Multivariate Normal Distribution"
author:
- "Pavel Krivitsky"
- "Deparment of Statistics"
- "University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---


# Packages

```{r, message=FALSE}
library(here)
library(mvtnorm)
library(GGally)
library(MVN)
library(dplyr)
library(readr)
```



# Challenges

Now, consider an extract from a dataset containing a number of chemical properties of Portuguese Vinho Verde wines and their quality, taken from the [UCI Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/Wine+Quality). Eleven properties of a wine were measured, along with a subjective quality score. This was done for red and for white wines. For the purposes of this exercise, we will focus on the following 7 variables:
* `fixed acidity`
* `volatile acidity`
* `free sulfur dioxide`
* `total sulfur dioxide`
* `density`
* `pH`
* `sulphates`

Let's explore the data and see if any variables would need to be transformed for multivariate analysis:
```{r}
wine <- read_csv(here("datasets","winequality-red.extract7.csv"))
```

### Task 1: Make a pairwise plot of this dataset and identify which variables may deviate from normality. Perform the hypothesis tests. Do the results agree?


### Task 2: Find an appropriate transformation for these variables, and check that they are now normal.*Note: There is no unique correct answer.


