---
title: "ZZSC5855 --- Multivariate Confidence Regions"
author:
- "Pavel Krivitsky"
- "based on Johnson and Wichern AMSA, Examples 5.3--5.4"
- "Department of Statistics"
- "University of New South Wales"
output:
  html_notebook:
    code_folding: "none"
  pdf_document: default
---

# Packages
```{r, message=FALSE}
library(here)
library(car) # For drawing ellipsoids.
library(GGally)
library(readr)
library(dplyr)
library(expm) # For %^% matrix power.
```


# Challenge
## One-sample intervals
We continue our analysis of of Portuguese red wines. We begin with the transformation suggested in the Multivariate Normality exercises:
```{r}
red.t <- read_csv(here("datasets","winequality-red.extract7.csv")) %>% mutate(`log sulphates`=log(`sulphates`), `root4 free sulfur dioxide` = (`free sulfur dioxide`)^(1/4), `root4 total sulfur dioxide` = (`total sulfur dioxide`)^(1/4), `log fixed acidity`=log(`fixed acidity`), `log volatile acidity`=log(`volatile acidity`), .keep="unused")
ggpairs(red.t)
```

### Task 1: Plot the data and the 95% confidence ellipse for the mean `log fixed acidity` and `log volatile acidity`.

#### Solution:
```{r}
plot(`log volatile acidity`~`log fixed acidity`, data=red.t, pch='.')
confidenceEllipse(lm(cbind(`log fixed acidity`, `log volatile acidity`)~1, data=red.t), add=TRUE)
```


### Task 2: Report the simultaneous confidence intervals for the mean `log fixed acidity` and `log volatile acidity` using the projections of the ellipsoid and the Bonferroni method. Which are wider? Why?

#### Solution:
```{r}
# From the demo above.
contrastCI <- function(l, n, m, S, CL=0.95){
    p <- length(m)
    # Note the use of "*c(-1,+1)" as plus-or-minus.
    c(crossprod(l,m)) + c(-1,+1)* c(sqrt(p*(n-1)/(n-p)*qf(CL, p, n-p) * t(l)%*%S%*%l/n))
}


# This constructs an n-vector with 1 in i'th position and 0s elsewhere.
e <- function(i, n) c(numeric(i-1), 1, numeric(n-i))

# Let's extract only the relevant variables from the dataset:
head(red.t12 <- select(red.t, `log fixed acidity`, `log volatile acidity`))

# Construct the summary statistics:
n <- nrow(red.t12)
p <- ncol(red.t12)
m <- colMeans(red.t12)
S <- cov(red.t12)

# Ellipsoid:
CCIs <- sapply(1:p, function(i) contrastCI(e(i,p), n, m, S))
colnames(CCIs) <- names(m)
# Bonferroni:
BCIs <- sapply(1:p, function(i) t.test(red.t12[[i]], conf.level = 1-(1-.95)/p)$conf.int)
colnames(BCIs) <- names(m)

CCIs
BCIs
```

We can even calculate the widths of the confidence intervals, i.e.,
```{r}
diff(CCIs)
diff(BCIs)
```

### Task 3: Now, report the simultaneous confidence intervals for all 7 variables in the dataset using the projections of the ellipsoid and the Bonferroni method. Which are wider? Why?

#### Solution:
```{r}
# Construct the summary statistics:
n <- nrow(red.t)
p <- ncol(red.t)
m <- colMeans(red.t)
S <- cov(red.t)

# Ellipsoid:
CCIs <- sapply(1:p, function(i) contrastCI(e(i,p), n, m, S))
colnames(CCIs) <- names(m)
# Bonferroni:
BCIs <- sapply(1:p, function(i) t.test(red.t[[i]], conf.level = 1-(1-.95)/p)$conf.int)
colnames(BCIs) <- names(m)

CCIs
BCIs

diff(CCIs)
diff(BCIs)
```

## Two-sample intervals and tests

Now, consider a related dataset with the same measurements of Portuguese *white* wines. We will apply the same transformation to those as for the red wines:
```{r}
white.t <- read_csv(here("datasets","winequality-white.extract7.csv")) %>% mutate(`log sulphates`=log(`sulphates`), `root4 free sulfur dioxide` = (`free sulfur dioxide`)^(1/4), `root4 total sulfur dioxide` = (`total sulfur dioxide`)^(1/4), `log fixed acidity`=log(`fixed acidity`), `log volatile acidity`=log(`volatile acidity`), .keep="unused")
ggpairs(white.t)
```

### Task 4: Produce a pairwise plot of the two datasets together, colour-coded by wine type. Comment on the apparent differences.

#### Solution:
```{r}
# This conveniently adds a "type" column indicating whether the wine is red or white.
head(wine.t <- bind_rows(list(red=red.t, white=white.t), .id="type"))
ggpairs(wine.t, mapping=aes(col=type, alpha=0.1), legend=1) + theme(legend.position="right")
```
* Density of white wines appears to be lower and more varied than that of the red wines.
* pH, sulphates, and volatile acidity of white wines appears to be lower than that of red wines while having about the same dispersion.
* Free sulphur dioxide and total sulphur dioxide in white wines appears to be higher than that of red wines and less dispersed.
* Fixed acidity of white wines is lower and less varied than that of the red wines.

Certain correlations differ. For example, fixed acidity and volatile acidity appear to be correlated for the red wines but not the white wines.


### Task 5: Assuming equal variances, implement the 2-Sample Hotelling's $T^2$ test described in **Comparison of two or more mean vectors** and use it to test the null hypothesis that all of the features of red and white wines are equal in expectation. Calculate the test statistic ($T^2$), the $\alpha=0.05$ critical value (i.e., rejection threshold) for $T^2$, and the $p$-value.

#### Solution:
```{r}
# Evaluate statistics
(p <- ncol(red.t))
(n.r <- nrow(red.t))
m.r <- colMeans(red.t)
S.r <- cov(red.t)

(n.w <- nrow(white.t))
m.w <- colMeans(white.t)
S.w <- cov(white.t)

# Pooled covariance matrix
S.p <- ((n.r-1)*S.r + (n.w-1)*S.w) / (n.r + n.w - 2)

# Test statistic
(T2 <- (m.r-m.w) %*% solve(S.p*(1/n.r+1/n.w)) %*% (m.r-m.w))

# Critical value:
qf(0.95, p, n.r+n.w-p-1) * (n.r+n.w-2)*p/(n.r+n.w-p-1)

# p-value:
Fstat <- T2 * (n.r+n.w-p-1)/(n.r+n.w-2)/p # The equivalent F-statistic.
pf(Fstat, p, n.r+n.w-p-1, lower.tail=FALSE)

# Note that although they produce the same p-values, these use different
# test statistics and parametrise them differently: some use T^2 directly,
# others scale it into F.

library(ICSNP)
HotellingsT2(red.t, white.t)

library(rrcov)
T2.test(red.t, white.t)

library(ergm)
approx.hotelling.diff.test(red.t, white.t, assume.indep=TRUE, var.equal=TRUE)
```


### Task 6: Repeat Task 5 assuming unequal variances. Also report the effective sample size for variance estimation ($\nu$).

#### Solution:
```{r}
# Evaluate statistics

# Test statistic
(T2 <- (m.r-m.w) %*% solve(S.r/n.r+S.w/n.w) %*% (m.r-m.w))

# nu
tr <- function(x) sum(diag(x))

(nu <- (p + p^2)/(
(tr((S.r%*%solve(S.r/n.r+S.w/n.w)/n.r)%^%2) + tr(S.r%*%solve(S.r/n.r+S.w/n.w)/n.r)^2)/n.r +
(tr((S.w%*%solve(S.r/n.r+S.w/n.w)/n.w)%^%2) + tr(S.w%*%solve(S.r/n.r+S.w/n.w)/n.w)^2)/n.w
))

# Critical value:
qf(0.95, p, nu-p+1) * nu*p/(nu-p+1)

# p-value:
Fstat <- T2 * (nu-p+1)/nu/p # The equivalent F-statistic.
pf(Fstat, p, nu-p+1, lower.tail=FALSE)

# Using the ergm package:
library(ergm)
approx.hotelling.diff.test(red.t, white.t, assume.indep=TRUE, var.equal=FALSE)
```

