---
title: "ZZSC5855 --- Multivariate Confidence Regions"
author:
- "Pavel Krivitsky"
- "based on Johnson and Wichern AMSA, Examples 5.3--5.4"
- "Department of Statistics"
- "University of New South Wales"
output:
  html_notebook:
    code_folding: "none"
  pdf_document: default
---


# Packages
```{r, message=FALSE}
library(here)
library(car) # For drawing ellipsoids.
library(GGally)
library(readr)
library(dplyr)
library(expm) # For %^% matrix power.
```

# The Microwave Ovens Example

We will use the data in Johnson and Wichern (2007), Examples 5.3--5.4 to illustrate Hotelling's $T^2$ confidence regions and to develop some intuition for their construction.

## Data

We begin by loading the data and transforming (See the Multivariate Normal exercises.):
```{r}
ovens4 <- read_csv(here("datasets","ovens.csv"))^(1/4)
ggpairs(ovens4)
```

## Summaries

Let's compute the basic data summaries:
```{r}
(n <- nrow(ovens4)) # sample size
(p <- ncol(ovens4)) # number of variables
(m <- colMeans(ovens4)) # sample mean
(S <- cov(ovens4)) # S
(Sinv <- solve(S)) # S^-1
```

## $T^2$ confidence ellipsoid

The following demonstration is intended to provide intuition for how a confidence ellipsoid is constructed.

We first define a function testing if a value is in the confidence ellipsoid. This is essentially a direct implementation of **Confidence region for mean vector**:
```{r}
in_ellipsoid <- function(x, n, m, Sinv, CL=0.95){
    p <- length(m)
    n*(m-x)%*%Sinv%*%(m-x) <= p*(n-1)/(n-p) * qf(CL, p, n-p)
}
```

We then draw a confidence ellipsoid in a very inefficient manner by taking a large number of points and testing them for being inside it. A more efficient way is to take the eigendecomposition of $\hat{\Sigma}$ and construct a parametric equation for its outline, then plot that. Let's draw both on the same plot.
```{r}
# Brute-force:
closeds <- seq(.50,.65,length.out=200)
opens <- seq(.55,.66,length.out=200)
xy <- expand.grid(closeds,opens)
z <- apply(xy, 1, in_ellipsoid, n, m, Sinv)
# Note: For meanings of asp=1 see ? par .
plot(xy[,1],xy[,2], col=z, xlab="closed",ylab="open", pch=".", asp=1)
points(m[1],m[2], col=2)

# Eigendecomposition:
(S.e <- eigen(S))

scl <- p*(n-1)/(n-p) * qf(0.95, p, n-p) / n
lines(rbind(m-S.e$vectors[,1]*sqrt(S.e$values[1]*scl), m+S.e$vectors[,1]*sqrt(S.e$values[1]*scl))) # Major axis
lines(rbind(m-S.e$vectors[,2]*sqrt(S.e$values[2]*scl), m+S.e$vectors[,2]*sqrt(S.e$values[2]*scl))) # Minor axis
```

The popular R package `car` provides a tool for drawing a confidence ellipsoid. Note the intermediate step of fitting a multivariate linear model with variables on the LHS:
```{r}
plot(open~closed, data=ovens4)
# I.e. fit a multivariate linear model (more on that later) with just the intercept:
confidenceEllipse(lm(cbind(closed,open)~1, data=ovens4), add=TRUE) # add=TRUE means don't start a new plot.
```

Any point in this ellipsoid is a plausible value for the joint population mean microwave radiation releases.

## Simultaneous intervals

In practice, we may also want to express the intervals numerically. We shall discuss two ways of doing so.

## Projection of the ellipsoid

Firstly, we can use "shadows" (projections) of the confidence ellipsoid onto the axes (or, technically, onto any other linear combination). The following function implements Theorem 4.3:
```{r}
contrastCI <- function(l, n, m, S, CL=0.95){
    p <- length(m)
    # Note the use of "*c(-1,+1)" as plus-or-minus.
    c(crossprod(l,m)) + c(-1,+1)* c(sqrt(p*(n-1)/(n-p)*qf(CL, p, n-p) * t(l)%*%S%*%l/n))
}
```
We can then evaluate these for unit vectors to get contrast CIs:
```{r}
l <- c(1,0) # Closed
(closedCI <- contrastCI(l, n, m, S))
l <- c(0,1) # Open
(openCI <- contrastCI(l, n, m, S))
```
Observe their relationship to the ellipsoid:
```{r}
confidenceEllipse(lm(cbind(closed,open)~1, data=ovens4), xlab="closed",ylab="open")
abline(v=closedCI)
abline(h=openCI)
```

## Bonferroni

Finally, Bonferroni: this is just a pair of one-sample $t$-intervals with at a higher individual CL:
```{r}
(closedBCI <- t.test(ovens4$closed,conf.level=.975)$conf.int)
(openBCI <- t.test(ovens4$open,conf.level=.975)$conf.int)
```

Observe that these are slightly narrower. This is because the ellipsoid intervals are simultaneously true for *every* linear combination, so we could, say, compute an interval for `open-closed` contrast using the ellipsoid projection without any additional "miss" probability.


