---
title: "ZZSC5855 --- R tutorial on basic multivariate summaries and visualisation"
author:
- "Pavel Krivitsky"
- "based on notes by Spiridon Penev"
- "Deparment of Statistics"
- "The University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---


The following example demonstrates basic multivariate summary statistics and graphics.

The "Iris" dataset is built into the `stats` package, so we can acquire it via the `data()` function:
```{r}
data(iris)
iris
```

# Basic data summaries

The following illustrates both built-in summary functions and `dplyr`.

To obtain basic 5-number summaries,
```{r}
summary(iris)
```

More generally, we can use the `apply` function to obtain univariate summaries for each variable:
```{r}
# Column means:
colMeans(iris[,-5])
# Column standard deviations:
apply(iris[-5], 2, sd)
# Column 95th %iles:
apply(iris[-5], 2, quantile, 0.95) # This evaluates: quantile(x, 0.95) for x each column in turn.
```

Similarly, we can calculate multivariate summaries:
```{r}
# Variance-covariance matrix
cov(iris[,-5])
# Correlation matrix
cor(iris[,-5])
```

We can also apply to subgroups. For this, we use the "pipe" operator `%>%` from package `magrittr`, which evaluates each function in turn, passing its results as the first argument of the next one (or `.`). We also use `map` function from the `purrr` package.
```{r}
library(magrittr)
library(purrr)
iris %>% # Start with dataset iris
  split(iris$Species) %>% # Split it according to its species column
  map(~.[-5]) %>% # For each species, throw away the last column (Species)
  map(cov)
```

# Visualisation
The most common visualisation for multivariate data is a pairwise plot: plot every variable against every other variable. A function that does it well---and automatically takes care of categorical variables---is `GGally`'s `ggpairs` function.

```{r}
library(GGally)
# aes() here is optional, telling it to colour-code in accodance with species, and using transparency (alpha).
ggpairs(iris, aes(colour=Species, alpha = 0.7))
```
