---
title: "ZZSC5855 --- Hypothesis Tests of Correlations"
author:
- "Pavel Krivitsky"
- "Department of Statistics"
- "University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---


# Packages
```{r, message=FALSE}
library(here)
library(readr)
library(dplyr)
library(GGally)
library(ggm)
```

# Intelligence, Weight, and Age
## Data
Recall the 3-variable example from before:
1. $X_{1}$: Intelligence
2. $X_{2}$: Weight
3. $X_{3}$: Age

Suppose that these correlations are actually based on sample size $n=20$:
$$\hat{\rho}=\begin{pmatrix}
1 & 0.6162 & 0.8267\\
0.6162 & 1 & 0.7321\\
0.8267 & 0.7321 & 1
\end{pmatrix}.$$

```{r, echo=FALSE}
R <- matrix(c(1     , 0.6162, 0.8267,
              0.6162, 1     , 0.7321,
              0.8267, 0.7321, 1     ),
            3,3)
```

## Testing correlations

We wish to test if there is sufficient evidence that weight and intelligence are correlated.

### Ordinary correlation

We can begin by using the $t$-test:
$$T=\hat{\rho}_{ij}\sqrt{\frac{n-2}{1-\hat{\rho}_{ij}^{2}}}=`r R[1,2]`\sqrt{\frac{n-2}{1-`r R[1,2]`^{2}}}=`r R[1,2]*sqrt((20-2)/(1-R[1,2]^2))`.$$
The $t$ distribution has $n-2$ degrees of freedom:
```{r}
R <- matrix(c(1     , 0.6162, 0.8267,
              0.6162, 1     , 0.7321,
              0.8267, 0.7321, 1     ),
            3,3)

2*pt(abs(R[1,2]*sqrt((20-2)/(1-R[1,2]^2))), 20-2, lower.tail=FALSE)
```
We can also use the `pcor.test()` function in `ggm`:
```{r}
pcor.test(R[1,2], 0, 20)
```

Alternatively, we can use Fisher's $Z$ approximation:
$$Z = \frac{1}{2}\log\frac{1+\hat{\rho}_{12}}{1-\hat{\rho}_{12}}=\frac{1}{2}\log\frac{1+`r R[1,2]`}{1-`r R[1,2]`}=`r log((1+R[1,2])/(1-R[1,2]))/2`.$$
Under the null hypothesis, this statistic will have mean 0 and variance $1/(n-3)$, so we compute a two-sided $p$-value as
```{r}
2*pnorm(abs(log((1+R[1,2])/(1-R[1,2]))/2), 0, sqrt(1/(20-3)), lower.tail=FALSE)
```

We conclude that the correlation is highly significant.

### Partial correlation

To test partial correlation given age, we compute the sample value,
```{r}
(r12.3 <- pcor(c(1,2,3), R))
```
Then,
$$Z = \frac{1}{2}\log\frac{1+\hat{\rho}_{12.3}}{1-\hat{\rho}_{12.3}}=\frac{1}{2}\log\frac{1+`r r12.3`}{1-`r r12.3`}=`r log((1+r12.3)/(1-r12.3))/2`,$$
and the variance is $1/(n-(p-r)-3)$. Here $p-r$ is the number of conditioned-on variables, which is in this case 1.
```{r}
2*pnorm(abs(log((1+r12.3)/(1-r12.3))/2), 0, sqrt(1/(20-1-3)), lower.tail=FALSE)
```

Similarly, we can use the $t$-test, using `pcor.test()`:
```{r}
pcor.test(r12.3, 1, 20)
```

We conclude that there is not sufficient evidence of correlation.

