---
title: "ZZSC5855 --- Hypothesis Tests of Correlations"
author:
- "Pavel Krivitsky"
- "Department of Statistics"
- "University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---


# Packages
```{r, message=FALSE}
library(here)
library(readr)
library(dplyr)
library(GGally)
library(ggm)
```


# Challenge
Recall the Portuguese red wines data:
```{r}
red.t <- read_csv(here("datasets","winequality-red.extract7.csv")) %>% mutate(`log sulphates`=log(`sulphates`), `root4 free sulfur dioxide` = (`free sulfur dioxide`)^(1/4), `root4 total sulfur dioxide` = (`total sulfur dioxide`)^(1/4), `log fixed acidity`=log(`fixed acidity`), `log volatile acidity`=log(`volatile acidity`), .keep="unused")
ggpairs(red.t)
```

Consider the variables `pH`, `log fixed acidity` and `log volatile acidity`.

### Task 1: Test the presence of correlation between pH and log volatile acidity, and interpret the results.

#### Solution:
```{r}
R <- cor(red.t)
pcor.test(R[2,7], 0, nrow(red.t))
```

There is sufficient evidence to believe that there is a correlation between pH and log volatile acidity.


### Task 2: Test the presence of partial correlation between pH and log volatile acidity given log fixed acidity, and interpret the results.

#### Solution:
```{r}
R <- cor(red.t)
pcor.test(pcor(c(2,7,6), R), 1, nrow(red.t))
```

There is not sufficient evidence to believe that there is a correlation between pH and log volatile acidity after controlling for log fixed acidity.

