---
title: "ZZSC5855 --- Discriminant Analysis"
author:
- Pavel Krivitsky
- Department of Statistics
- University of New South Wales
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---


# Packages
```{r}
library(here)
library(readr)
library(MASS) # lda(), qda()
library(dplyr) # Also, we want select() from here and not MASS, so we load it second.
library(GGally)
library(heplots) # boxM
```


# Challenges

Recall the dataset `pizza.csv`, containing nutritional data from a variety of pizza brands:

`brand`
: Pizza brand (class label)

`id`
: Sample analysed

`prot`
: Amount of protein per 100 grams in the sample

`fat`
: Amount of fat per 100 grams in the sample

`ash`
: Amount of ash per 100 grams in the sample

`sodium`
: Amount of sodium per 100 grams in the sample

`carb`
: Amount of carbohydrates per 100 grams in the sample

`cal`
: Amount of calories per 100 grams in the sample

```{r}
pizza <- read_csv(here("datasets","pizza.csv"))
ggpairs(pizza, mapping=aes(col=brand, alpha=0.3))
```

### Task 1: Compare LDA and QDA for classifying the pizza brand based on measurements using in-sample predictions (i.e., not cross-validated). Which looks better?
Hint: the error rate is `1-sum(diag(CM))` where `CM` is the confusion matrix.


### Task 2: Now compare LDA and QDA using cross-validation. How does your answer change?


### Task 3: Use your selected classifier to classify a pizza that has the following properties. Obtain predicted probabilities.

* protein content of 12
* fat content of 15
* ash content of 1.5
* sodium content of 0.6
* carbohydrate content of 25
* caloric content of 3

**Hint:** See `?predict.lda`.

