---
title: "ZZSC5855--- Support Vector Machines"
author:
- "Pavel Krivitsky"
- "Department of Statistics"
- "University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---


# Packages
```{r}
library(here)
library(GGally)
library(e1071) # svm()
library(pROC) # ROC curves
library(dplyr)
library(readr)
```


# Challenges
Recall the dataset `pizza.csv`, containing nutritional data from a variety of pizza brands:

`brand`
: Pizza brand (class label)

`id`
: Sample analysed

`prot`
: Amount of protein per 100 grams in the sample

`fat`
: Amount of fat per 100 grams in the sample

`ash`
: Amount of ash per 100 grams in the sample

`sodium`
: Amount of sodium per 100 grams in the sample

`carb`
: Amount of carbohydrates per 100 grams in the sample

`cal`
: Amount of calories per 100 grams in the sample

```{r}
pizza <- read_csv(here("datasets","pizza.csv")) %>% mutate(brand=factor(brand))
ggpairs(pizza, mapping=aes(col=brand, alpha=0.3))
```

### Task 1: Select and tune an SVM for classifying the pizza brand based on measurements. What is its cross-validated accuracy?


### Task 2: Use your selected classifier to classify a pizza that has the following properties.

* protein content of 12
* fat content of 15
* ash content of 1.5
* sodium content of 0.6
* carbohydrate content of 25
* caloric content of 3



