---
title: "ZZSC5855 --- Canonical Correlation Examples"
author:
- "Pavel Krivitsky"
- "Department of Statistics"
- "University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---

```{r, setup2, echo=FALSE, cache=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=8, message=FALSE, warning=FALSE, error=FALSE)
```


# Packages
```{r, message=FALSE}
library(here)
library(CCA)
library(CCP)
library(GGally)
library(dplyr)
library(readr)
library(purrr)
select <- dplyr::select # Just to be safe.
```

# Fitness Club Example

Three physiological and three exercise variables were measured on twenty middle aged men in a fitness club. Canonical correlation is used to determine if the physiological variables are related in any way to the exercise variables. 

## Data

```{r}
'weight waist pulse chins situps jumps
   191  36  50   5  162   60
   189  37  52   2  110   60
   193  38  58  12  101  101
   162  35  62  12  105   37
   189  35  46  13  155   58
   182  36  56   4  101   42
   211  38  56   8  101   38
   167  34  60   6  125   40
   176  31  74  15  200   40
   154  33  56  17  251  250
   169  34  50  17  120   38
   166  33  52  13  210  115
   154  34  64  14  215  105
   247  46  50   1   50   50
   193  36  46   6   70   31
   202  37  62  12  210  120
   176  37  54   4   60   25
   157  32  52  11  230   80
   156  33  54  15  225   73
   138  33  68   2  110   43' %>%
textConnection() %>% read.table(header=TRUE) -> fitness

fitness

```
## Exploratory data analysis

```{r}
summary(fitness)
ggpairs(fitness)
```
The data do have some outliers.

## Canonical Correlation Analysis
### Estimation and visualisation
Fitting canonical correlations is straightforward:
```{r}
X <- fitness[1:3]
Y <- fitness[4:6]
print(fitness.cc <- cc(X,Y))
```
What do we see? Firstly,

`cor`
: Canonical correlations for the first, second, and third pair of canonical vectors.

`xcoef`, `ycoef`
: Canonical coefficients for the first and second set of variables. (Variables are in rows, and vectors are in columns.)

`xscores`, `yscores`
: Canonical scores for each observation, produced by taking the inner product of each observation vector and a canonical coefficient vector. These are the scores whose correlations we calculate.

`corr.X.xscores`, `corr.Y.xscores`, etc.
: These are the correlations between the canonical scores and the original variables. Note that these may or may not be the same as the coefficients: in particular, if there is correlation within $X$ or within $Y$, some of the variables may be "redundant".

We can also produce plots. It helps to set `var.label=TRUE` to label which variables have which canonical correlation weights:
```{r}
plt.cc(fitness.cc, var.label=TRUE)
```
Here, the left-hand-side plot plots the `corr.X.xscores` for the first two canonical vectors for each of the variables and similarly for the `corr.Y.xscores`. Then, the horizontal position of a variable reflects its relationship with the first canonical vector and its vertical position the second. In this case, we might conclude that in the most important dimension of correlation waist and weight, as they increase, are negatively correlated with chins and situps.

### Standardised canonical correlation coefficients
To compare coefficients between variables *in the same group*, it can be helpful to standardise the variables first. This can be done by the `scale` function.
```{r}
Xs <- scale(X)
Ys <- scale(Y)
print(fitness.ccs <- cc(Xs,Ys))
```
### Hypothesis testing (asymptotic)
Hypothesis testing only requires the correlations, sample sizes, and variable counts. A number of tests are possible:
```{r}
n <- nrow(fitness)
p <- ncol(X)
q <- ncol(Y)
p.asym(fitness.cc$cor, n, p, q, tstat = "Wilks")
p.asym(fitness.cc$cor, n, p, q, tstat = "Hotelling")
p.asym(fitness.cc$cor, n, p, q, tstat = "Pillai")
p.asym(fitness.cc$cor, n, p, q, tstat = "Roy")
```
Here, the rows "$i$ to $j$" should be read as "$H_0: \rho_i = \dotsb = \rho_j = 0$, where $\rho_k$ is the $k$th canonical correlation."

We thus see that the first canonical correlation does not appear to meet significance at the conventional $\alpha=0.05$ for most of the tests. The fact that the data deviates from normality and the relatively small sample size probably explain the differences between the different tests.

Where available, the first test tests for significance of the first canonical correlation (i.e., any linear dependence between datasets), the second for the significance of the second (i.e., there are linear relationships between the datasets beyond the first), etc.

### Permutation testing
Permutation tests can only test for the first canonical correlation, but they are valid even if the data are not even close to MVN:
```{r}
p.perm(X,Y, type = "Wilks")
p.perm(X,Y, type = "Hotelling")
p.perm(X,Y, type = "Pillai")
p.perm(X,Y, type = "Roy")
```

# When only correlations are available: Chicken data

Based on Example 10.4, p. 552 in Johnston and Wichern, this is a study of canonical correlations between leg and head bone measurements: $X_{1},X_{2}$ are skull length and skull breadth, respectively; $X_{3},X_{4}$ are leg bone measurements: femur and tibia length, respectively. Observations have been taken on $n=276$ White Leghorn chicken. The example is chosen to also illustrate how a canonical correlation analysis can be performed when the original data are not given but the empirical correlation matrix (or empirical covariance matrix) is available.

## Data

```{r}
" 1.0      .505   .569   .602
   .505   1.0     .422   .467
   .569   .422   1.0     .926
   .602   .467   .926   1.0" %>% 
   textConnection() %>% scan() %>%
   matrix(4,4, byrow=TRUE) -> chicken
   
rownames(chicken) <- colnames(chicken) <- c("head1","head2","leg1","leg2")

n <- 276

chicken
```

## Canonical Correlation Analysis

### Estimation
Here, we are using the algorithm from Section 7.4.

```{r}
matpow <- function(A, p){
with(eigen(A), vectors%*%diag(values^p,nrow(A))%*%t(vectors))
}
(ccdecomp <- eigen(matpow(chicken[3:4,3:4],-1/2)%*%chicken[3:4,1:2]%*%solve(chicken[1:2,1:2])%*%chicken[1:2,3:4]%*%matpow(chicken[3:4,3:4],-1/2)))
(ccors <- sqrt(ccdecomp$values))
(b <- matpow(chicken[3:4,3:4],-1/2) %*% ccdecomp$vectors[,1])
(a <- 1/sqrt(ccdecomp$values[1])*solve(chicken[1:2,1:2])%*%chicken[1:2,3:4]%*%b)
```

### Hypothesis testing (asymptotic)
```{r}
library(CCP)
p <- 2
q <- 2
p.asym(ccors, n, p, q, tstat = "Wilks")
p.asym(ccors, n, p, q, tstat = "Hotelling")
p.asym(ccors, n, p, q, tstat = "Pillai")
p.asym(ccors, n, p, q, tstat = "Roy")
```

