---
title: "ZZSC5855 --- Canonical Correlation Examples"
author:
- "Pavel Krivitsky"
- "Department of Statistics"
- "University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---

```{r, setup2, echo=FALSE, cache=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=8, message=FALSE, warning=FALSE, error=FALSE)
```


# Packages
```{r, message=FALSE}
library(here)
library(CCA)
library(CCP)
library(GGally)
library(dplyr)
library(readr)
library(purrr)
select <- dplyr::select # Just to be safe.
```


# Challenge
Recall the dataset `pizza.csv`, containing nutritional data from a variety of pizza brands:

`brand`
: Pizza brand (class label)

`id`
: Sample analysed

`prot`
: Amount of protein per 100 grams in the sample

`fat`
: Amount of fat per 100 grams in the sample

`ash`
: Amount of ash per 100 grams in the sample

`sodium`
: Amount of sodium per 100 grams in the sample

`carb`
: Amount of carbohydrates per 100 grams in the sample

`cal`
: Amount of calories per 100 grams in the sample


We will focus on brand A:
```{r}
pizzaA <- read_csv(here("datasets","pizza.csv")) %>% filter(brand=="A") %>% select(-brand)
```

### Task 1: Calculate the canonical correlations between caloric, ash, and sodium content (set 1) and fat, protein, and carbohydrate content (set 2). What do the vectors represent?


### Task 2: Test the significance of correlation between the two sets; comment on the results. How many canonical correlations are significant, and what do they represent?

