---
title: "ZZSC5855 --- Partial Correlations"
author:
- "Pavel Krivitsky"
- "Department of Statistics"
- "University of New South Wales"
output:
  html_notebook: default
  pdf_document: default
---


# Packages

```{r, message=FALSE}
library(here)
library(readr)
library(dplyr)
library(GGally)
library(ggm)
```

# Intelligence, Weight, and Age

Suppose that three variables have been measured for a set of schoolchildren:

1. $X_{1}$: Intelligence
2. $X_{2}$: Weight
3. $X_{3}$: Age

The number of observations was large enough so that one can assume the empirical correlation matrix $\hat{\rho}\in\mathcal{M}_{3,3}$ to be the true correlation matrix:
$$\hat{\rho}=\begin{pmatrix}
1 & 0.6162 & 0.8267\\
0.6162 & 1 & 0.7321\\
0.8267 & 0.7321 & 1
\end{pmatrix}.$$
This suggests there is a high degree of positive dependence between weight and intelligence. But let's compute the partial correlation adjusting for age: $$\hat{\rho}_{12.3}=\frac{\hat{\rho}_{12}-\hat{\rho}_{13}\hat{\rho}_{23}}{\sqrt{(1-\hat{\rho}_{13}^2)(1-\hat{\rho}_{23}^2)}}=\frac{0.6162-0.8267\times 0.7321}{\sqrt{(1-0.8267^2)(1-0.7321^2)}}=0.0286.$$
Alternatively, R can do it for us:
```{r}
R <- matrix(c(1     , 0.6162, 0.8267,
              0.6162, 1     , 0.7321,
              0.8267, 0.7321, 1     ),
            3,3)
# Calculate the partial correlations for each pair of variables given the rest, and take the one between X1 and X2:
parcor(R)[1,2]

# Calculate the partial correlation between variables 1 and 2 given variable 3 in R:
pcor(c(1:2,3), R)
```

Either way, after the effect of age is adjusted for, there is virtually no correlation between weight and intelligence, i.e. weight obviously plays little part in explaining intelligence.

We will also use these data to study *testing* correlations.

